---
title: "FORMATIVE ASSESSMENT 2"
author: "Gabriel Tagaytay"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1. Use R to illustrate that the probability of getting:**  

**- a head is 0.5 if a fair coin is tossed repeatedly;**  

* **Simulate Coin Tosses**
```{r}
set.seed(123)
n_trials <- 10000

coin_tosses <- sample(c("Head", "Tail"), size = n_trials, replace = TRUE)
```

* **Calculate the probability of getting heads**  
```{r}
prob_heads <- sum(coin_tosses == "Head") / n_trials
print(paste("Probability of getting a Head:", prob_heads))
```  

**- a red card is 0.5 if cards are drawn repeatedly with replacement from a well-shuffled deck;**  

* **Simulate Drawing Cards**
```{r}
suits <- rep(c("Red", "Black"), each = 26)
n_trials <- 10000

draws <- sample(suits, size = n_trials, replace = TRUE)
``` 

* **Calculate probability of drawing a red card**  
```{r}
prob_red_card <- sum(draws == "Red") / n_trials
print(paste("Probability of drawing a Red Card:", prob_red_card))
```  

**- an even number is 0.5 if a fair die is rolled repeatedly.**  

* **Simulate Die Roll**  
```{r}
n_trials <- 10000
die_rolls <- sample(1:6, size = n_trials, replace = TRUE)
```

* **Calculate probability of rolling an even number**  
```{r}
prob_even <- sum(die_rolls %% 2 == 0) / n_trials
print(paste("Probability of rolling an Even Number:", prob_even))
```  



**2. An experiment consists of tossing two fair coins. Use R to simulate this experiment 100 times and obtain the relative frequency of each possible outcome. Hence, estimate the probability of getting one head and one tail in any order.**  

* **Set seed for reproducibility**  
```{r}
set.seed(123)
```

* **Number of simulations**
```{r}
n_simulations <- 100
```

* **Simulate tossing two coins 100 times**
```{r}
coin1 <- sample(c("H", "T"), size = n_simulations, replace = TRUE)
coin2 <- sample(c("H", "T"), size = n_simulations, replace = TRUE)
```

* **Combine outcomes**
```{r}
outcomes <- paste(coin1, coin2, sep = "")
```

* **View first few outcomes**  
```{r}
head(outcomes)
```
* **Calculate relative frequency of each outcome**  
```{r}
outcome_freq <- table(outcomes) / n_simulations
print(outcome_freq)
```
* **Estimate the probability of getting one head and one tail (HT or TH)**  
```{r}
prob_one_head_one_tail <- sum(outcome_freq[c("HT", "TH")])
print(paste("Estimated probability of getting one head and one tail:", prob_one_head_one_tail))
```





[CLICK HERE TO VIEW THE R MARKDOWN FILEON GITHUB!!!](https://github.com/gabriyeeel/Formattive-Assessments-Files/blob/main/FA2.Rmd)
