---
title: "FA3"
author: "Gabriel Tagaytay"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 2: Binary Communication Channel
## Given probabilities
```{r}
P_T0 <- 0.7  
P_T1 <- 0.3
P_R0_given_T0 <- 0.95
P_R1_given_T1 <- 0.75
P_R1_given_T0 <- 1 - P_R0_given_T0
P_R0_given_T1 <- 1 - P_R1_given_T1
```

* **(a) Probability that a 1 was received**
```{r}
P_R1 <- (P_R1_given_T0 * P_T0) + (P_R1_given_T1 * P_T1)
print(paste("P(R1) =", P_R1))
```

* **(b) Probability that a 1 was transmitted given that a 1 was received**
```{r}
P_T1_given_R1 <- (P_R1_given_T1 * P_T1) / P_R1
print(paste("P(T1 | R1) =", P_T1_given_R1))
```


# Problem 7: Employee Errors
## Given probabilities
```{r}
P_J <- 0.1
P_A <- 0.3
P_V <- 0.6
P_E_given_J <- 0.08
P_E_given_A <- 0.05
P_E_given_V <- 0.01
```

* **(a) Overall probability of an error**
```{r}
P_E <- (P_E_given_J * P_J) + (P_E_given_A * P_A) + (P_E_given_V * P_V)
print(paste("P(E) =", P_E))
```

* **(b) Probability of each person making an error given that an error occurred**
```{r}
P_J_given_E <- (P_E_given_J * P_J) / P_E
P_A_given_E <- (P_E_given_A * P_A) / P_E
P_V_given_E <- (P_E_given_V * P_V) / P_E

print(paste("P(J | E) =", P_J_given_E))
print(paste("P(A | E) =", P_A_given_E))
print(paste("P(V | E) =", P_V_given_E))
```

## Identifying the most likely person to have written an erroneous program
```{r}
most_likely <- which.max(c(P_J_given_E, P_A_given_E, P_V_given_E))
most_likely_person <- c("Jane", "Amy", "Ava")[most_likely]
print(paste("Most likely person to have written the erroneous program:", most_likely_person))
```
[CLICK HERE TO VIEW THE R MARKDOWN FILE ON GITHUB!!!](https://github.com/gabriyeeel/Formattive-Assessments-Files/blob/main/FA6.Rmd)