---
title: "FORMATIVE ASSESSMENT 4"
author: "Gabriel Tagaytay"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 5: A geospatial analysis system has four sensors supplying images. The percent-age of images supplied by each sensor and the percentage of images relevant to a query are shown in the following table.   

# Given data
```{r}
data <- data.frame(
  Sensor = 1:4,
  Percentage_Images_Supplied = c(15, 20, 25, 40),
  Percentage_Relevant_Images = c(50, 60, 80, 85)
)
```

# Convert percentages to decimal form
```{r}
data$Percentage_Images_Supplied <- data$Percentage_Images_Supplied / 100
data$Percentage_Relevant_Images <- data$Percentage_Relevant_Images / 100
```

# Compute overall relevant percentage
```{r}
overall_relevance <- sum(data$Percentage_Images_Supplied * data$Percentage_Relevant_Images)
```

# Convert back to percentage form
```{r}
overall_relevance <- overall_relevance * 100
```

# Print result
```{r}
cat("Overall Percentage of Relevant Images:", overall_relevance, "%\n")
```  

# Problem 6: # Problem 6: A fair coin is tossed twice. 
Let \(E_1\) be the event that both tosses have the same outcome, that is, \(E_1 = \{HH, TT\}\).  
Let \(E_2\) be the event that the first toss is a head, that is, \(E_2 = \{HH, HT\}\).  
Let \(E_3\) be the event that the second toss is a head, that is, \(E_3 = \{TH, HH\}\).  
Show that \(E_1\), \(E_2\), and \(E_3\) are pairwise independent but not mutually independent.
```{r}
prob_HH <- 1/4
prob_HT <- 1/4
prob_TH <- 1/4
prob_TT <- 1/4
```

# Define events
```{r}
prob_E1 <- prob_HH + prob_TT  # Both tosses are the same
prob_E2 <- prob_HH + prob_HT  # First toss is a head
prob_E3 <- prob_HH + prob_TH  # Second toss is a head
```

# Compute intersections
```{r}
prob_E1_E2 <- prob_HH
prob_E1_E3 <- prob_HH
prob_E2_E3 <- prob_HH
prob_E1_E2_E3 <- prob_HH
```

# Check pairwise independence
```{r}
independent_E1_E2 <- (prob_E1_E2 == prob_E1 * prob_E2)
independent_E1_E3 <- (prob_E1_E3 == prob_E1 * prob_E3)
independent_E2_E3 <- (prob_E2_E3 == prob_E2 * prob_E3)
```

# Check mutual independence
```{r}
mutual_independence <- (prob_E1_E2_E3 == prob_E1 * prob_E2 * prob_E3)
```

# Print results
```{r}
cat("Pairwise Independence:\n")
cat("E1 and E2:", independent_E1_E2, "\n")
cat("E1 and E3:", independent_E1_E3, "\n")
cat("E2 and E3:", independent_E2_E3, "\n")

cat("\nMutual Independence:\n")
cat("E1, E2, and E3:", mutual_independence, "\n")
```