---
title: "FA5"
author: "Gabriel Tagaytay"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)  # Load ggplot2 for visualization
library(tidyr)    # Load tidyr for reshaping data
```
# PROBLEM NO. 6  
**Given data**  
```{r}
messages <- c(0.40, 0.25, 0.35)  # Percentage of messages
errors <- c(0.01, 0.02, 0.015)   # Percentage of errors
```
**Create a data frame for visualization**
```{r}
server_data <- data.frame(
  Server = c("Server 1", "Server 2", "Server 3"),
  Percentage_Messages = messages * 100,  # Convert to percentage
  Percentage_Errors = errors * 100       # Convert to percentage
)
```
**Visualization**
```{r}
server_data_long <- pivot_longer(server_data, cols = -Server, 
                                 names_to = "Category", values_to = "Percentage")
```
**Create bar plot**
```{r}
ggplot(server_data_long, aes(x = Server, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Messages and Errors per Server",
       x = "Server",
       y = "Percentage",
       fill = "Category") +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", "red"))
```

* **(a) Probability of receiving an email containing an error**  
```{r}
P_E <- sum(messages * errors)
cat("Probability of receiving an email containing an error:", P_E, "\n")
```
* **(b) Probability that a message will arrive without error**
```{r}
P_no_E <- 1 - P_E
cat("Probability that a message will arrive without error:", P_no_E, "\n")
```
* **(c) Given that a message arrives without error, probability it was sent through Server 1**
```{r}
P_no_E_given_S1 <- 1 - errors[1]  # Probability of no error given Server 1
P_S1_given_no_E <- (P_no_E_given_S1 * messages[1]) / P_no_E
cat("Probability that the message was sent through Server 1 given no error:", P_S1_given_no_E, "\n")
```  

# PROBLEM NO. 9  
**Given probabilities**
```{r}
P_A <- 0.20
P_B <- 0.70
P_C <- 0.10

P_G_given_A <- 0.10
P_G_given_B <- 0.40
P_G_given_C <- 0.20
```

**(a) Probability of not upgrading**
```{r}
P_G <- (P_A * P_G_given_A) + (P_B * P_G_given_B) + (P_C * P_G_given_C)
P_Gc <- 1 - P_G
cat("P(G^c):", P_Gc, "\n")
```

**(b) Posterior probability P(B|G)**
```{r}
P_B_given_G <- (P_B * P_G_given_B) / P_G
cat("P(B|G):", P_B_given_G, "\n")
```

**(c) Other probabilities**
```{r}
P_Gc_given_B <- 1 - P_G_given_B
P_B_given_Gc <- (P_B * P_Gc_given_B) / P_Gc

P_C_given_G <- (P_C * P_G_given_C) / P_G
P_Gc_given_C <- 1 - P_G_given_C
P_C_given_Gc <- (P_C * P_Gc_given_C) / P_Gc

cat("P(B|G^c):", P_B_given_Gc, "\n")
cat("P(C|G):", P_C_given_G, "\n")
cat("P(C|G^c):", P_C_given_Gc, "\n")
```  
# TREE DIAGRAM  
**Load necessary libraries**
```{r}
library(DiagrammeR)
library(webshot2)
library(DiagrammeRsvg)
library(rsvg)
```

**Define probabilities**
```{r}
P_A <- 0.20   # Probability of not buying (A)
P_B <- 0.70   # Probability of buying (B)
P_C <- 0.10   # Probability of being undecided (C)

P_G_given_A <- 0.10  # Given A, probability of upgrading
P_G_given_B <- 0.40  # Given B, probability of upgrading
P_G_given_C <- 0.20  # Given C, probability of upgrading
```

**Manually define the GraphViz diagram**
```{r}
graph <- grViz("
digraph {
  Survey -> A [label='20%']
  Survey -> B [label='70%']
  Survey -> C [label='10%']

  A -> G_A [label='10%']
  A -> NG_A [label='90%']
  
  B -> G_B [label='40%']
  B -> NG_B [label='60%']
  
  C -> G_C [label='20%']
  C -> NG_C [label='80%']
}
")
```

**Save as PDF**
```{r}
svg_file <- tempfile(fileext = ".svg")  # Temporary SVG file
pdf_file <- "tree_diagram.pdf"          # Output PDF file
```

**Convert GraphViz diagram to SVG, then to PDF**
```{r}
export_svg(graph) %>% charToRaw() %>% rsvg_pdf(pdf_file)
```
**Insert the PDF into the RMarkdown document**
```{r}
knitr::include_graphics("tree_diagram.pdf")
```  

# PROBLEM NO. 13  
**Given probabilities**
```{r}
P_Internet <- 0.70  # Probability that the spyware comes through the Internet
P_Email <- 0.30     # Probability that the spyware comes through Email

P_Detected_given_Internet <- 0.60  # Probability that spyware is detected if it came through Internet
P_Detected_given_Email <- 0.80     # Probability that spyware is detected if it came through Email
```

* **(a) What is the probability that this spyware infects the system?**
```{r}
P_NotDetected_given_Internet <- 1 - P_Detected_given_Internet  # 0.40
P_NotDetected_given_Email <- 1 - P_Detected_given_Email        # 0.20

P_Infection <- (P_Internet * P_NotDetected_given_Internet) + 
               (P_Email * P_NotDetected_given_Email)

cat("Probability that the spyware infects the system:", P_Infection, "\n")
```

* **(b) If the spyware is detected, what is the probability that it came through the Internet?**  
```{r}
P_Detected <- (P_Internet * P_Detected_given_Internet) + 
              (P_Email * P_Detected_given_Email)

P_Internet_given_Detected <- (P_Internet * P_Detected_given_Internet) / P_Detected

cat("Probability that the spyware came through the Internet given that it was detected:", P_Internet_given_Detected, "\n")
```
[CLICK HERE TO VIEW THE R MARKDOWN FILE ON GITHUB!!!](https://github.com/gabriyeeel/Formattive-Assessments-Files/blob/main/FA5.Rmd)
